---
title: "Reproducible Research: Project 1, Activity Data"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r, echo=T}
data <- read.csv(file="activity.csv", colClasses = c("numeric","character", "numeric"))
head(data)
data$date <- as.Date(data$date,"%Y-%m-%d")
```

## Histogram and Mean total number of steps taken per day
```{r, echo=T}
dailySteps <- with(data,aggregate(steps,by=list(date=data$date), sum,na.rm = T))
AvgDailySteps <- mean(dailySteps$x)
MedianDailySteps <- median(dailySteps$x)
hist(dailySteps$x, main = "Histogram total number of daily steps", xlab = "Total number of daily steps", breaks =10)
```

The median number of steps taken daily is: `r as.integer(MedianDailySteps)`

The mean number of steps taken daily is: `r as.integer(AvgDailySteps)`

## The average daily activity pattern
```{r, echo=T}
dailyPattern <- with(data,aggregate(steps,by=list(interval=data$interval), mean,na.rm = T))
maxInterval=dailyPattern$interval[which.max(dailyPattern$x)]
dailyPattern$interval=dailyPattern$interval/100
with(dailyPattern, plot(interval, x, xlab= "interval(min)", ylab="Avg steps", type="l"))
```

Maximum steps interval is: `r floor(maxInterval/100)` : `r maxInterval%%100` 


## Imputing missing values

Total Number of rows that have NA data: `r sum(complete.cases(data)==0)`
All the NA data is in steps column.
Imputing the data:
```{r, echo=T}
NumNARows <- sum(complete.cases(data)==0)
library(mice)
md.pattern(data, plot=T)
imputedData <- mice(data, method="pmm",m=1)
completeData <- complete(imputedData,1)
```

Calculate average and median daily steps and show the histogram:

```{r, echo=T}
dailyStepsC <- with(completeData,aggregate(steps,by=list(date=data$date), sum,na.rm = T))
AvgDailyStepsC <- mean(dailyStepsC$x)
MedianDailyStepsC <- median(dailyStepsC$x)

hist(dailyStepsC$x, xlab = "Total number of daily steps", breaks =10, main = "Histogram daily steps, imputed data")
```


The median number of steps taken daily (imputed) is: `r as.integer(MedianDailyStepsC)`

The mean number of steps taken daily (imputed) is: `r as.integer(AvgDailyStepsC)`

The median is the same as the original data, but the mean has changenged slightly.


## Are there differences in activity patterns between weekdays and weekends?

Check which dates are a weekday and ones are weeked days:
```{r, echo=T}
isweekend <- as.factor(weekdays(completeData$date) %in% c("Saturday","Sunday"))
newData <- cbind(completeData,isweekend)
avg_weekday <- aggregate(newData$steps,by=list(isweekend=newData$isweekend, interval=newData$interval),mean)
avg_weekday$wday <-factor(avg_weekday$isweekend,labels = c("weekday","weekend"))
```

Plot the average number of steps for weed days and weekend days.

```{r, echo=T}
library(ggplot2)
g <- ggplot(avg_weekday,aes(interval,x))
g+geom_line()+facet_grid(avg_weekday$wday~.)+labs(y="Avg Steps")
```